MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-12-10 chapters/MetaTracts.tex
roll back 1 commit

\section{Introduction and Motivation}
\label{sec:intro}
Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components. The material of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of these continuously rising demands. 
%CFRPs have been successfully introduced in aeronautic and automotive %applications within recent years, replacing parts made of conventional %materials such as aluminum or steel. But the application areas of CFRPs %are not limited on these two industries. For a wide range of different %areas ranging from marine and civil engineering, wind energy, to sporting %goods manufacturing, CFRP is highly interesting. 
%The main reason therefore is found in CFRP materials having desirable %characteristics such as high specific stiffness, high specific strength %and high corrosion resistance. 
CFRP materials are increasingly replacing conventional materials such as aluminum or steel, as they have desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance. 
Moreover, CFRP materials show these characteristics at considerably lower weight. At the same time, highly complex and integrated components, which were previously impossible to manufacture may be produced from CFRPs. Primary structures and highly loaded components in aeronautics are one example.

Typically, carbon fiber reinforced polymer components and more specifically CFRP laminates with endless carbon fibers as addressed in this work, consist of two main components:
a matrix, which acts as a bonding component, and the reinforcements, which allow for achieving the desired characteristics. Various production processes are used to manufacture CFRP laminates. Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth are also referred to as fabric and form the geometrical structure of the final CFRP materials. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth in similar or different orientation. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the properties of the CFRP laminate. Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of the matrix is to act as a bonding between the individual carbon fiber bundles. After curing, the production process of the CFRP laminate is finished.

The increasing share of CFRPs as well as the complexity of material system and final components has generated a strong demand towards non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. Ultrasonic testing (UT) is the most commonly used method for this purpose. While UT provides a quick and cost-efficient overview of the material, it also shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component. Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. In contrast to UT, XCT generates a highly detailed 3D volumetric representation of the scanned specimen. In cone beam XCT geometry the specimen is placed on a rotary table between X-ray source and detector. The X-rays passing through the specimen get attenuated by the materials present. By transferring the X-rays in a scintillator layer into visible light, the detector records the corresponding 2D attenuation image (penetration image). The specimen is rotated stepwise recording at each position a 2D attenuation image to generate the full series of attenuation images of a $360^\circ$ rotation for a complete reconstruction of the data volume ~\cite{heinzl-2008-thesis}. Finally the series of penetration images is reconstructed to a 3D volumetric dataset of the specimen. 
While cone beam XCT can reach voxel sizes below 500 $nm$ generating high resolution volume data for comprehensive and detailed analyses, unfortunately there is still a trade-off between viewport and image resolution. The magnification reached within an XCT scan is determined by the used distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: while higher resolutions decrease the viewport but show more details, lower resolutions allow for larger viewports and thus larger portions of the specimen.

@@

Figure~\ref{fig:data-char} provides an example of our targeted dataset type. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
Our work is motivated from the recent progress in two interrelated fields: Firstly, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}. Secondly, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}. 

@@

We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. It provides answers to questions such as: Are these bundles in contact at a particular position in the data? What is the relative orientations of the contacting bundles?

Providing answers to the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We present and evaluate a method which uses visualization techniques to gain insight into our data.
We interpret and advance techniques from diffusion tensor imaging to extract and visualize geometric structures from XCT data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
\begin{figure}[tb]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=\linewidth]{images_pvis/workflow.pdf}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
\end{figure}
Figure~\ref{fig:flowchart} shows an over view of our approach. Starting from the XCT data, we first generate MetaTracts. Then we cluster the MetaTracts to generate and visualize fiber bundles. For a detailed description of the MetaTracts generation process see Sec.~\ref{sec:approach}. For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}. Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.

@@

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
In DTI, a similarity measure based on factors such as proximity between fibers is used to group fiber tracts into bundles. Extensive research has been done on automatic DTI fiber clustering methods \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}. 
Clustering requires choosing a suitable proximity measure and choosing a method for grouping ``close" fibers.
Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}. Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers that are close for certain distance and then diverge, are clustered separately. Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors. The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering~\cite{Moberts2005, Zhang2008} and those using spectral clustering~\cite{ Brun2004,jonasson2005, ODonnell2007}.

@@

Lomov et al.~\cite{Lomov2010, Lomov2001} and Straumit et al.~\cite{Straumit2014} in recent papers, discussed the problems and current available solutions in geometric modeling of three dimensional composites. Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position? The second step builds the geometry of the model, answering questions relating to placement of bundles in space their orientations and dimensions.

%Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes have received extensive research. Recently, similar techniques have been applied to fiber tracing in composite materials. 
To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers.

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
\includegraphics[width=\linewidth]{images_pvis/data-char.pdf}
\caption{Data characteristics: (a) Rendering of the data. (b) 2D slices along the X-axis and (c) 2D slice along the Z-axis. (d) Magnified region marked in green square (c). Multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}
Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles. Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$. The data type is unsigned 8 bit integer. Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}a shows a volume rendering of the dataset. Figure~\ref{fig:data-char}b shows 2D slices along the $X$- and Figure~\ref{fig:data-char}c in $Z$-axis. Figure~\ref{fig:data-char}d shows the magnified version of the green region of interest. 
The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve in the 2D slice images. Figure~\ref{fig:data-char}d contains two bundles going in opposite directions. The separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles may differ in terms of the amount of fibers in the bundle. Figure~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern, the fiber bundles cross each other in different orientations. In general, the number of orientations is limited by the weaving pattern. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles. 
We make the following \textit{assumptions} on our data:
\begin{itemize}[noitemsep]
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item Local orientation: each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item Connectivity: moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

@@

The input to the first stage of our pipeline, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$. As output, each grid location becomes associated with an unit vector representing the local orientation and a real value [0,1] that represents a measure of the reliability of the local orientation. We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel.
The Hessian matrix represents the local curvature. The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image.

The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.

Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular. They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.  In order to determine reliable Hessians, we compute 
\color{red}{ the same metric} \color{blue}{a similar metric}\color{black}. We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details. Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude. Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix. Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$). The sign indicates if the vessel is bright in a dark background or dark in a bright background. In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$). The following measures are defined in~\cite{Frangi1998}.  

@@

In Equation~\ref{RB}, $\mathcal{R_{B}}$ provides a measure of deviation from a blob like structure while in Equation~\ref{RA}, $\mathcal{R_{A}}$ distinguishes between plate-like and line-like structure. Grey-scale variations and close proximity of the fibers in our data make the Hessians computed at each voxels susceptible to errors. Thus, we compute ``reliable Hessians" ($R_H$) to determine which locations in the volume provide reliable local orientation. ($R_H$ is similar to the ``vesselness" in~\cite{Frangi1998}).
%Thus, we compute the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.

@@

We note some intrinsic differences between the DTI and our XCT data. Fiber traces can be created in DTI using a standard fiber tracking algorithm following the principal direction of diffusion using a fourth order Runge-Kutta method~\cite{Brun2003}. The principal direction based on the Hessian matrix works best when the tubular structures are well separated from the surrounding. This is not the case for our data. The local orientaion at each voxel is inherently more noisy.

@@

\subsection {MetaTracts properties and description} 
\label{subsec:mtprop}
%\begin{figure*}
%  \centering
%  \subfloat[Meta tract generation in 2D]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1b.eps}}
%  \subfloat[Orientation based similarity measure]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1c.eps}}
%  \caption{Simplified examples of meta tract generation process and orientation based similarity measure}
%  \label{fig:algo}
%\end{figure*}
%\begin{figure*}
%  \centering
%  \subfloat[]{\includegraphics[width=0.2\textwidth]{imagesMetaTracts/algo-1d}}
%  %\subfloat[Orientation based similarity measure]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1c.eps}}
%  \caption{Simplified example of meta tract generation process.}
%  \label{fig:algo}
%\end{figure*}
The input to this step is a grid in $\mathbb{R}^3$ where each grid point is associated with a normalized local orientation and a corresponding $R_H$ value.
Traditional integral curve based techniques cannot be used to extract fiber bundle traces from the reliable Hessians because of the spurious nature of the Hessian based local orientation. Thus instead of building fiber traces, we wish to find an abstract representation of the fibers. We build on the key assumptions on the data, namely ``local orientation" and ``connectivity" while taking into account the noise and lack of resolution. We do this by interpreting the underlying geometric structure of the fibers as a set of cylinders.

 MetaTacts  are a coarse and simple approximation of integral curves in the form of a continuous chain of cylindrical tubes in $\mathbb{R}^3$ traversing the fiber bundles embedded in the input data. MetaTracts share the following \textit{properties}:

\begin{enumerate}[noitemsep]
\item{Each MetaTract is associated with a continuous set of cylinders.}
\item{Each MetaTract is associated with a start point which is a grid vertex.}
\item{Each individual cylinder in a MetaTract has a constant length, a constant radius and a start point (which is also a grid vertex).}
\item{Each cylinder in a MetaTract (except for the first one) is connected to the previous cylinder at its start point.}
\item{Each cylinder is parallel to the local orientation at its start point.}
\end{enumerate}

@@

The output of the above step is a set of MetaTracts. The next step is to cluster the MetaTracts into bundles. 
Carbon fibers have both orientation and geometric proximity information, both of which can be used for clustering. Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with fibers which partially overlap (Figure~\ref{fig:length_distribution}b).  
Instead of creating a heuristic which artificially combines the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
We found that different clustering techniques performed preferably for different measures (Sec.~\ref{subsec:clus_choice}).  The two step clustering also made the problem more tractable and  the parameters intuitive for the end users. For the orientation clustering, we use dimension reduction followed by K-means clustering. We then use hierarchical clustering to further subdivide each cluster based on the geometric proximity measures.

@@

 %glass
We tested our method on three datasets with varying characteristics. We report results on two here and third in the supplemental section.  
Dataset 1 is described in Sec.~\ref{sec:char_data} (Figure~\ref{fig:data-char}). Dataset 1 has eleven separate fiber bundles (5 along one orientation and 6 along the other). All of them were identified correctly by MetaTracts. With different cross-section sizes and varying degree of curvature of the bundles, dataset 1 is a complicated dataset.
Figure~\ref{fig:orientation_clustering} and Figure~\ref{fig:crop-16-decomp} shows the result of dataset 1 decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters followed by voxelization and mesh extraction. Note how the thin and curved purple cluster bundle 7 is extracted well (Figure~\ref{fig:orientation_clustering}e, Figure~\ref{fig:crop-16-decomp}d).

>>>>

\section{Introduction and Motivation}
\label{sec:intro}
Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components. The material of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands. CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing parts made of conventional materials such as aluminum or steel. But the application areas of CFRPs are not limited on these two industries. For a wide range of different areas ranging from marine and civil engineering, wind energy, to sporting goods manufacturing, CFRP is highly interesting. The main reason therefore is found in CFRP materials having desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance. CFRP materials show these characteristics at considerably lower weight compared to conventional materials. At the same time, highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs. Primary structures and highly loaded components in aeronautics are one example.

Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates we are addressing in this work, consist of two main components:
a matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.  Various production processes are used to manufacture CFRP laminates with endless carbon fibers. Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also referred to as fabric. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the properties of the CFRP laminate. Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of this matrix component is to act as a bonding between the individual carbon fiber bundles. After curing, the production process of the CFRP laminate is finished.

The increasing share of CFRPs as well as the complexity of material system and final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers. On one hand, ultrasonic testing provides a quick and cost-efficient overview of the material. On the other hand, it shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component. Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen. In the typical scanning geometry of industrial XCT, the specimen is placed on a rotary table between X-ray source and detector. When passing through the specimen the X-rays are attenuated by the material in the specimen. The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light. To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}. Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen. 
In state of the art devices, XCT can reach voxel sizes below 500 $nm$. XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry. The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but show more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

@@

Figure~\ref{fig:data-char} provides an example of our targeted dataset. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
Our work is motivated from the recent progress in two interrelated fields: On one hand, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}. On the other hand, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}. 

@@

 We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. For example, spatial context answers questions such as: Are these bundles in contact at a particular position in the data? What is the relative orientations of the contacting bundles?

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data.
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
\begin{figure}[tb]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=\linewidth]{images_pvis/workflow.pdf}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
\end{figure}
Figure~\ref{fig:flowchart} shows an over view of our approach. Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters. For a detailed description of the MetaTracts generation process, see Sec.~\ref{subsec:MetaTracts-generation}. For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}. Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.

@@

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
In DTI, a similarity measure based on factors such as proximity between fibers is used to group fiber tracts into bundles. Extensive research has been done on automatic DTI fiber clustering methods~ \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}. 
Clustering requires choosing a suitable proximity measure and choosing a method for grouping ``close" fibers.
Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}. Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers that are close for certain distance and then diverge, are clustered separately. Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors. The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering~\cite{Moberts2005, Zhang2008} and those using spectral clustering~\cite{ Brun2004,jonasson2005, ODonnell2007}.

@@

Lomov et al.~\cite{Lomov2010} in a recent paper, discussed the current available solutions in geometric modeling of three dimensional composites. Modeling of the composites starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position? The second builds the geometry of the model, answering questions relating to placement of bundles in space their orientations and dimensions.

%Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes have received extensive research. Recently, similar techniques have been applied to fiber tracing in composite materials. 
To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers, which is the main scope of this work.

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
\includegraphics[width=\linewidth]{images_pvis/data-char.pdf}
\caption{Data characteristics: (a) Rendering of the data. (b) 2D slices along the X-axis and (c) 2D slice along the Z-axis. (d) Magnified region marked in green square (c). Multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}
Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles. Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$. The data type is unsigned 8 bit integer. Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}a shows a volume rendering of the dataset. Figure~\ref{fig:data-char}b shows 2D slices along the $X$- and Figure~\ref{fig:data-char}c in $Z$-axis. Figure~\ref{fig:data-char}d shows the magnified version of the green region of interest. 
The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve in the 2D slice images. Figure~\ref{fig:data-char}d contains two bundles going in opposite directions. The separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles may differ in terms of the amount of fibers in the bundle. Figure~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern, the fiber bundles cross each other in different orientations. In general, the number of orientations is limited by the weaving pattern. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles. 
We make the following assumptions on our data:
\begin{itemize}[noitemsep]
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item Local orientation: each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item Connectivity: moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

@@

The input to the first stage of our pipeline, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$. As output, each grid location becomes associated with a vector representing the local orientation and a real value [0,1] that represents a measure of the reliability of the local orientation. We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel. The principal directions, in which the local second order structure of the image can be decomposed, are given by the eigen decomposition of the Hessian matrix which describes the local curvature.
The eigenvector corresponding to the smallest eigenvalue gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.

Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular. They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.  In order to determine reliable Hessians, we compute a similar metric. We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details. Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude. Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix. Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$). The sign indicates if the vessel is bright in a dark background or dark in a bright background. In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$). The following measures are defined in~\cite{Frangi1998}.  

@@

In Equation~\ref{RB}, $\mathcal{R_{B}}$ provides a measure of deviation from a blob like structure while in Equation~\ref{RA}, $\mathcal{R_{A}}$ distinguishes between plate-like and line-like structure. Grey-scale variations and close proximity of the fibers in our data make the Hessians computed at each voxels susceptible to errors. Thus, we compute the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.

@@

We note some intrinsic differences between the DTI and our XCT data. Fiber traces can be created in DTI using a standard fiber tracking algorithm following the principal direction of diffusion using a fourth order Runge-Kutta method~\cite{Brun2003}. The principal direction based on the Hessian matrix works best when the tubular structures are well separated from the surrounding. This is not the case for our data. The local orientaion at each voxel is inherently more noisy. Thus, techniques popular in DTI cannot be directly applied to XCT data.

@@

\subsection {MetaTracts properties and description} 
\label{subsec:mtprop}
%\begin{figure*}
%  \centering
%  \subfloat[Meta tract generation in 2D]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1b.eps}}
%  \subfloat[Orientation based similarity measure]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1c.eps}}
%  \caption{Simplified examples of meta tract generation process and orientation based similarity measure}
%  \label{fig:algo}
%\end{figure*}
%\begin{figure*}
%  \centering
%  \subfloat[]{\includegraphics[width=0.2\textwidth]{imagesMetaTracts/algo-1d}}
%  %\subfloat[Orientation based similarity measure]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1c.eps}}
%  \caption{Simplified example of meta tract generation process.}
%  \label{fig:algo}
%\end{figure*}
The input to this step is a grid in $\mathbb{R}^3$ where each grid point is associated with a normalized local orientation and a corresponding $R_H$ value.
Traditional integral curve based techniques cannot be used to extract fiber bundle traces from the reliable Hessians because of the spurious nature of the Hessian based local orientation. Thus instead of building fiber traces, we wish to find an abstract representation of the fibers. We build on the key assumptions on the data, namely ``local orientation" and ``connectivity" while taking into account the noise and lack of resolution. We do this by interpreting the underlying geometric structure of the fibers as a set of polylines (cylinders). MetaTacts  are a coarse and simple approximation of integral curves in the form of a continuous chain of cylindrical tubes in $\mathbb{R}^3$ traversing the fiber bundles embedded in the input data. MetaTracts share the following \textit{properties}:

\begin{enumerate}[noitemsep]
\item{Each MetaTract is associated with a continuous set of cylinders.}
\item{Each MetaTract is associated with a start point which is a grid vertex.}
\item{Each individual cylinder in a MetaTract has a constant length, a constant radius and a start point (which is also a grid vertex).}
\item{Each cylinder in a MetaTract (except for the first one) is connected to the previous cylinder at its start point.}
\item{Each cylinder is locally parallel to the local orientation at its start point.}
\end{enumerate}

@@

The output of the above step is a set of MetaTracts. The next step is to cluster the MetaTracts into bundles. 
Carbon fibers have both orientation and geometric proximity information, both of which can be used for clustering. Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with fibers which partially overlap (Figure~\ref{fig:length_distribution}b).  
Instead of creating a heuristic which artificially combines the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
We found that different clustering techniques performed preferably for different measures.  The two step clustering also made the problem more tractable and  the parameters intuitive for the end users. For the orientation clustering, we use dimension reduction followed by K-means clustering. We then use hierarchical clustering to further subdivide each cluster based on the geometric proximity measures.

@@

 %glass
We tested our method on three datasets with varying characteristics. We report results on two due to lack of space.  
Dataset 1 is described in Sec.~\ref{sec:char_data} (Figure~\ref{fig:data-char}). Dataset 1 has eleven separate fiber bundles (5 along one orientation and 6 along the other). All of them were identified correctly by MetaTracts. With different cross-section sizes and varying degree of curvature of the bundles, dataset 1 is a complicated dataset.
Figure~\ref{fig:orientation_clustering} and Figure~\ref{fig:crop-16-decomp} shows the result of dataset 1 decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters followed by voxelization mesh extraction. MetaTracts correctly identified the fiber bundles. Note how the thin and curved purple cluster bundle 7 is extracted well (Figure~\ref{fig:orientation_clustering}e, Figure~\ref{fig:crop-16-decomp}d).
